An Evolutionary Algorithm for the Multiobjective Risk-Equity Constrained Routing Problem

### 2011-11-30 trunk/DAW_hazmatHypE/trunk/paper/hazmatRouting_EvolOptim.tex
rewrote intro (including some new references that have to be put into the BibteX file)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The transportation of hazardous materials ({\em hazmat} from now on) has received a large interest in recent years, which results from the increase in public awareness of the dangers of hazmats and the enormous amount of hazmats being transported \cite{cd2008a}. The main target of this problem is to select routes from a given origin-destination pair of nodes such that the risk for the surrounding population and the environment is minimized---without producing excessive economic costs. 
%
When solving such a problem by minimizing both cost and the total risk, typically several vehicles share the same (short) routes which results in high risks associated to regions surrounding these paths whereas other regions are not affected. In this case, one may wish to distribute the risk in an equitable way over the population and the environment. Several studies consider this additional minimization of the equity risk, but most of them consist of a \emph{single} origin-destination hazmat routing for a specific hazmat, transport mode and vehicle type (see for example \cite{aeb2000a, cd2008a}).
% 
% 
% and can be classified into the \textit{resolution-equity-based methods} and the \textit{model-equity-based methods}. In resolution-equity-based methods, the equity constraints are taken into account only indirectly via a dissimilarity index between paths which is to be maximized \cite{JOH92, LOM93, KUB97, aeb2000a}. Model-equity-based methods, on the other hand, formalize the risk equity as constraints directly within the model. In \cite{GOP90a, GOP90b}, for example, the authors propose an equity shortest path model that minimizes the total risk of travel while the difference between the risks imposed on any two arbitrary zones does not exceed a given threshold.
%
A more realistic \emph{multi}-commodity flow model was proposed in \citep{cd2008a} where each commodity is considered as one hazmat type. The objective function is formulated as the sum of the economical cost and the cost related to the consequences of an incident for each material. To deal with risk equity, the costs are defined as functions of the flow traversing the arcs which imposes an increase of the arc's cost and risk when the number of vehicles transporting a given material increases on the arc.

The majority of all hazmat routing studies deal with a single-objective scenario although the problem itself is multiobjective in nature and it is important to study the trade-offs among the objectives. Evolutionary Multiobjective Optimization (EMO) algorithms are able to compute a set of solutions showing these trade-offs within a single algorithm run which is the reason why we propose to use them for the problem of hazmat routing in this study (Sec.~\ref{S_MEO}). Before, we formalize the routing of hazmat problem with three objectives (minimize total routing cost, total routing risk {\em and} risk equity) as a multicommodity flow model as in \cite{cd2008a} since this model is the most realistic one permitting to manage several hazmat types simultaneously (Sec.~\ref{S_RECRP}).

% see for example \cite{cd2008a, cgr2007, gkbk1990, za2004}, but all problem formulations include the equity risk as constraints and not as an additional objective. \citet{gian1998} on the other hand considers the equity risk as separate objective but uses a weighted sum to aggregate his four objectives into a single-objective problem. The main reason why, to the best of our knowledge, no study considered a fully multiobjective model with the equity risk as objective might be the resulting non-linearity of the problem which make classical optimization algorithms such as branch-and-cut \TODON{equity is an objective in \cite{{gian1998}} ... We solve the broblem with a branch and price algorithm because the model is a multicommodity flow model, and this model is the more adapted one for this problem because it permit to manage simultaneously many hazmat types. The min-max formulation of the objective is easily linearized. I give you the major axes of the work: (1) Solve the presented problem (optimize cost, risk and equity) (2) what are models and resolution methods proposed in the litterature (3) what is the model retained and why (in our case, the multicommodity model) (4) how to handle equity with this model ? (5) how to solve it ? (branch and price is adapted for these models, but can not hadle many objectives $->$ evolutionnary algorithms) } infeasible and led us to our choice of proposing an evolutionary multiobjective algorithm as heuristic here.

%The computation of routes with a fairly distributed risk consists in generating dissimilar origin-destination paths, i.e paths which relatively don't impact the same zones. Solution approaches

%Section~\ref{S_RECRP} defines the problem of transporting hazardous materials with equity risk as a three-objective optimization problem with the minimization of the total routing cost, of the total routing risk {\em and} of the risk equity as the objectives---the latter broadly defined as the risk shared by a set of regions that compose the geographical area under consideration. Based on this formulation, Sec.~\ref{S_MEO} presents an evolutionary multiobjective optimization algorithm before Sec.~\ref{S_FW} concludes the abstract.

%In \cite{cd2008a} was proposed a multi-commodity flow model for this problem, where each commodity is considered as one hazmat type. To deal with risk equity, the costs are defined as functions of the flow traversing the arcs, this imposes an increase of the arc's cost and risk when the number of vehicles transporting a given material increases on the arc. The major drawback of this model is that it can lead to the saturation of the arcs of shortest paths and, therefore, does not guarantee the equity target.

%In \cite{BEL10}, the authors consider a similar model but with additional equity constraints and objectives. The objective is the minimization of the maximum risk equity imposed on the population and the environment and the model is solved using a pranch-and-price algorithm.

%As hazmat routing is multiobjective in nature, we introduce in this work a multiobjective version of this problem. Our goal is the minimization of the total routing cost, of the total routing risk {\em and} of the risk equity, the latter broadly defined as the risk shared by a set of regions that compose the geographical area under consideration.

%This paper is organized as follows. We describe the problem in section \ref{S_RECRP}, we give its mathematical formulation in section \ref{S_AOM}, in section \ref{S_MEO} we show how multiobjective evolutionary methods can be applied to this problem and in section \ref{S_FW} we exopse our further work.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the remainder of this paper, we investigate how the above proposed evolutionary multiobjective optimization algorithm can cope with realistic instances of the multiobjective risk-equity constrained routing problem. The experimental analyses have two main goals. On the one hand, we show that the algorithm is efficient and can find multiple non-dominated solutions on practical examples in reasonable time. On the other hand, we investigate which of the two proposed initialization schemes is favourable and when this is the case. In addition, a few more experiments are conducted to show that the standard parameters of the algorithm behave as with other test problems. To this end, we vary the population size of the algorithm as well as its environmental selection scheme.

@@

Overall, it is to say that the approach is very general and can easily be adapted if the problem formulationi is (slightly) changed as it is often the case in practice, e.g., if different numbers of trucks are available. Also in case of different truck loads, it is only the objective function code that needs to be changed---a big advantage of blackbox optimizers such as the proposed EMO algorithm.

@@

In future work, one might want to look further into adding (simple) heuristics within the EMO algorithm in order to improve the convergence of the algorithm. For example, one can observe that the solutions produced by the algorithm sometimes uses one and the same arc within a truck path, hence introducing cylces. With the three objectives cost, total and equity risk, such cycles can be easily prevented within the variation operator which might eventually result in a faster convergence of the algorithm to the Pareto front.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The transportation of hazardous materials ({\em hazmat} from now on) has received a large interest in recent years, which results from the increase in public awareness of the dangers of hazmats and the enormous amount of hazmats being transported \citep{etv2007a,cd2008a}. The main target of this problem is to select routes from a given origin-destination pair of nodes such that the risk for the surrounding population and the environment is minimized---without producing excessive economic costs.

More specifically, one can distinguish between two main problems of different scale \citep{etv2007a}. The \emph{local} hazmat routing problem aims at finding the route for a single shipment from a source to a destination node in a given network; the \emph{global} hazmat routing problem, on the other hand, looks as the macro level of shipping several hazmat types simultaneously from multiple sources to multiple destinations---with much less studies on the latter \citep{etc2007a}. 
When solving such a global hazmat routing problem by minimizing both cost and the total risk, typically several vehicles share the same (short) routes which results in high risks associated to regions surrounding these paths whereas other regions might not be affected at all. In this case, one may wish to distribute the risk in an equitable way over the population and the environment. Although several studies consider this additional minimization of the equity risk, most of them restrict the model to a \emph{single} origin-destination hazmat routing for a specific hazmat, transport mode and vehicle type (see for example \cite{aeb2000a}).
More realistic \emph{multi}-commodity flow model have been proposed where each commodity is considered as one hazmat type \citep{cd2008a} but the majority of all hazmat routing studies do not consider the multi-objective nature of the hazmat routing problem but optimize only the risk \cite{gkbk1990a} or an aggregation or scalarization of the objective functions instead \citep{cd2008a}. An exception is \citep{dgs2005a} where the focus lies on finding a set of dissimilar paths that are Pareto-optimal with respect to cost and risk instead of solving the multiobjective hazmat routing problem directly. Moreover, only one origin-destination pair is considered.

As it is important for a decision maker to study the trade-offs among the objectives, we treat the global hazmat routing problem of finding a set of routes minimizing total costs, total risk, and equity risk, from a multiobjective perspective. Evolutionary Multiobjective Optimization (EMO) algorithms have been shown to be able to compute a set of solutions showing the trade-offs between multiple objectives within a single algorithm run in particular when other (exact) algorithms have difficulties to approach the Pareto front, e.g, if the front is too large. As general black-box algorithms, they also allow to cope with changes in the problem formulation which occurs frequently in real-world applications.

In this study, we therefore propose an EMO algorithm for the global hazmat routing problem (Sec.~\ref{S_MEO}). Before, we formalize the specific multiobjective risk-equity constrained routing problem with three objectives (minimize total routing cost, total routing risk {\em and} risk equity) as a multicommodity flow model as in \cite{cd2008a} since this model is the most realistic one permitting to manage several hazmat types simultaneously (Sec.~\ref{S_RECRP}). Section~\ref{sec:experiments} is then devoted to the experimental analysis of this algorithm and its application to realistic instances in the Lazio region of Italy before Sec.~\ref{S_FW} concludes the paper.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental analysis}\label{sec:experiments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the remainder of this paper, we investigate how the above proposed evolutionary multiobjective optimization algorithm can cope with realistic instances of the multiobjective risk-equity constrained routing problem. The experimental analyses have two main goals. On the one hand, we show that the algorithm is efficient and can find multiple non-dominated solutions on practical examples in reasonable time. On the other hand, we investigate which of the two proposed initialization schemes is favourable and when this is the case. In addition, a few more experiments are conducted to show that the standard parameters of the algorithm behave as with other test problems. To this end, we vary the population size of the algorithm as well as its environmental selection scheme.

@@

Overall, it is to say that the approach is very general and can easily be adapted if the problem formulation is (slightly) changed as it is often the case in practice, e.g., if different numbers of trucks or different types of vehicles with different loads are available. Also in the case when the definition of total and equity risk changes, it is only the objective function code that needs to be changed---a big advantage of blackbox optimizers such as for the proposed EMO algorithm.

@@

In future work, one might want to look further into adding (simple) heuristics within the EMO algorithm in order to improve the convergence of the algorithm. For example, one can observe that the solutions produced by the algorithm sometimes uses one and the same arc within a truck path, hence introducing cylces. With the three objectives cost, total and equity risk, such cycles can be easily prevented within the variation operator which might eventually result in a faster convergence of the algorithm to the Pareto front. Another path for future research is the consideration of more realistic problem instances. Stochastic risks or dynamically changing instances (for example due to road closures or traffic jams) would be interesting to address---challenges that have been dealt with by EMO algorithms in other application domains already before \cite{deb2001,cvl2007a}.
